Inferring probabilistic stellar rotation periods using Gaussian processes

### 2017-02-06 documents/ms.tex

The uncertainties produced by the GP method are more representative than those
produced by the ACF and periodogram methods because the posterior PDF of the
period is explored using MCMC.
Whilst they are an improvement, unfortunately these uncertainties are still
underestimated in the majority cases.
Of the rotation periods recovered from the simulated light curves, only 25\%
of the measured periods lie within 1 $\sigma$ of the true period.
50\% lie within 2 $\sigma$ and 66\% within 3 $\sigma$.
\tdmcomment{What are these numbers for the uninformed-prior experiment?}
The largest outlier is 114 $\sigma$ away from the true value.
In the majority of these cases the uncertainties are underestimated due to the
multi-modal nature of the period posterior PDFs, which makes them difficult to
sample. \tdmcomment{Is there any evidence of this?  Doesn't sound convincing to me.}
The underestimated uncertainties can also be attributed to the model choice.
Although a GP model is more appropriate than a sinusoid, it is still an
imperfect, effective model.
A perfect physical model of the star would still produce better results with
more representative uncertainties.

>>>>

The uncertainties produced by the GP method are more representative than those
produced by the ACF and periodogram methods because the posterior PDF of the
period is explored using MCMC.
Whilst they are an improvement, unfortunately these uncertainties are still
underestimated in the majority cases.
Of the rotation periods recovered from the simulated light curves, only 25\%
of the measured periods lie within 1 $\sigma$ of the true period.
50\% lie within 2 $\sigma$ and 66\% within 3 $\sigma$.
\tdmcomment{What are these numbers for the uninformed-prior experiment?}
The largest outlier is 114 $\sigma$ away from the true value.
In the majority of these cases the uncertainties are underestimated due to the
multi-modal nature of the period posterior PDFs, which makes them difficult to
sample. \tdmcomment{Is there any evidence of this?  Doesn't sound convincing to me.
I would think the multi-modality would only be relevant if the posterior is 
closer to the 1/2x or 2x period, but this isn't most of them.}
The underestimated uncertainties can also be attributed to the model choice.
Although a GP model is more appropriate than a sinusoid, it is still an
imperfect, effective model.
A perfect physical model of the star would still produce better results with
more representative uncertainties. \tdmcomment{Not sure saying ``a perfect
model would give better results'' is worth saying...}
